---
title: "Philip Morgan Consulting Research Briefing"
subtitle: "Managing remote developers"
author: "Philip Morgan"
date: November 12, 2019 - v1
fontsize: 13pt
urlcolor: blue
always_allow_html: true
output:
  tufte::tufte_html:
    toc: true
    toc_depth: 2
  tufte::tufte_handout:
    toc: true
    toc_depth: 2
---

\newpage
The question at the heart of this study is simple: How large is the remote team you manage, why did you choose to work with remote developers, and what's difficult about it?

# Key Findings

1. If your intuition about why companies work with remote developers is "cost and talent availability", then your intuition is correct.
2. The median remote team size for this study's sample group is 7.
3. The primary difficulties in working with remote teams are communication and time zone overlap.

\newpage
# All Findings

This research used a short, focused questionnaire. Below are the summarized responses to each question.

## Q: "In the last year, have you recommended working remotely to others?"

Of the 45 respondents who had managed dev team that is overseas OR a remote/distributed team that has just 2 or 3 hours/day of time zone overlap in the last year, 86.7% recommended it to others.

## Q: How many people are part of the team you manage?

```{r echo = FALSE, warning=FALSE, fig.fullwidth=TRUE}
suppressMessages(library(data.table))
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
suppressMessages(library(ggplot2))
suppressMessages(library(plotly))
suppressMessages(library(formattable))


sizetable<-fread("teamsize.csv",sep=",")
sizetable$Percentage <- (sizetable$Respondents / 45)
sizetable$Percentage <- percent(sizetable$Percentage)

kable(sizetable, caption = "Remote team sizes", col.names = c("Remote team size", "Respondents managing a team of this size", "Percentage")) %>%
    row_spec(0, bold = T)

p <- plot_ly(sizetable, x = ~Size, y = ~Respondents)
p <- layout(p, xaxis = list(type = "log"), title="Remote team sizes")
tmpFile <- tempfile(fileext = ".png")
suppressMessages(export(p, file = tmpFile))
```

\newpage
## Q: What is the #1 reason your company chose to hire remote developers?

I standardized the free-form answers to this question using a simple coding system. The tabulated coded responses are below. Note that some respondents provided more than one reason for choosing to hire remote developers.

```{r echo = FALSE, warning=FALSE}
reasonstable<-fread("reasons.csv",sep=",")
reasonstable$Percentage <- (reasonstable$Respondents / 45)
reasonstable$Percentage <- percent(reasonstable$Percentage)


kable(reasonstable, caption = "Reasons for remote", col.names = c("Reason for remote team", "Respondents indicating this reason", "Percentage")) %>%
    row_spec(0, bold = T)
```

## Q: What seems to be more difficult when managing remote developers?

Tabulated coded responses are below. Note that some respondents listed more than one difficulty in managing remote developers.

```{r echo = FALSE, warning=FALSE}
difficultiestable<-fread("difficulty.csv",sep=",")
difficultiestable$Percentage <- (difficultiestable$Respondents / 45)
difficultiestable$Percentage <- percent(difficultiestable$Percentage)

kable(difficultiestable, caption = "Difficulties with remote", col.names = c("Primary difficulty", "Respondents indicating this difficulty", "Percentage")) %>%
    row_spec(0, bold = T)
```

# Insights

The median remote team size is 7. 

**Team building** is a difficulty that disproportionately effects large teams (9 or larger remote team members).

**Dealing with time zone overlap** is a difficulty that disproportionately effects small teams (7 or fewer remote team members).

**Access to better and cheaper talent** is the broadly shared set of reasons that both large and small teams choose to work with remote developers. ^[I'm aware of how painfully obvious this finding is, but I call it out anyway because there's some quantitative data to support it.] Some small teams work with remote developers because they wanted to preserve an existing relationship through some kind of relocation.

# Method

In September 2019, I surveyed a group of Tech Leads and VP's of Engineering from Computer Software, Internet, and Information Technology and Services companies. This group was a convenience opt-in sample recruited using LinkedIn. I screened respondents by asking "In the last year, have you managed a dev team that is overseas OR a remote/distributed team that has just 2 or 3 hours/day of time zone overlap with you?"

- 56 people from LinkedIn completed the survey. 45 (80%) moved beyond the screener question.

# Open data

You can inspect and use for your own purposes the anonymized survey responses generated by this research: <https://docs.google.com/spreadsheets/d/1uKwJddwN0RKlkAb58Qldmc8xkVDtqHzQH4xmuk8NHXM/edit?usp=sharing>

# Discussion

I welcome discussion with you about the conclusions you might draw from this data, especially if they differ from mine. 

# Brought To You By Philip Morgan Consulting

I'm Philip Morgan, and I'm focused on helping dev shops cultivate, commercialize, and monetize expertise. 

There's an underlying tension inherent in this expertise project: what dev shops need to do to cultivate exceptionally valuable expertise is often in direct conflict with what they need to do to serve clients -- and generate revenue -- today. My work helps resolve this tension. 

Learn more: 
<https://philipmorganconsulting.com/research/>

