---
title: "Philip Morgan Consulting Research Briefing"
subtitle: "Managing remote developers"
author: "Philip Morgan"
date: November 6, 2019 - v0.8
abstract: |
  ABSTRACT TODO
fontsize: 13pt
urlcolor: blue
always_allow_html: true
output:
  tufte::tufte_handout:
    toc: true
    toc_depth: 2
  tufte::tufte_html:
    toc: true
    toc_depth: 2
---

\newpage
The question at the heart of this study is simple: How large is the remote team you manage, why did you choose to work with remote developers, and what's difficult about it?

# Key Findings

1. F1
2. F2
3. F3
4. F4

\newpage
# All Findings

This research used a short, focused questionnaire. Below are the summarized responses to each question.

## Q: "In the last year, have you recommended working remotely to others?"

Of the 45 respondents who had managed dev team that is overseas OR a remote/distributed team that has just 2 or 3 hours/day of time zone overlap in the last year, 86.7% recommended it to others.

## Q: How many people are part of the team you manage?

```{r echo = FALSE, warning=FALSE, fig.fullwidth=TRUE}
suppressMessages(library(data.table))
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
suppressMessages(library(ggplot2))
suppressMessages(library(plotly))
suppressMessages(library(formattable))


sizetable<-fread("teamsize.csv",sep=",")
sizetable$Percentage <- (sizetable$Respondents / 45)
sizetable$Percentage <- percent(sizetable$Percentage)

kable(sizetable, caption = "Remote team sizes", col.names = c("Remote team size", "Respondents managing a team of this size", "Percentage")) %>%
    row_spec(0, bold = T)

p <- plot_ly(sizetable, x = ~Size, y = ~Respondents, type="bar")
p <- layout(p, xaxis = list(type = "log"), title="Remote team sizes")
tmpFile <- tempfile(fileext = ".png")
suppressMessages(export(p, file = tmpFile))
```

\newpage
## Q: What is the #1 reason your company chose to hire remote developers?

I standardized the free-form answers to this question using a simple coding system. The tabulated coded responses are below. Note that some respondents provided more than one reason for choosing to hire remote developers.

```{r echo = FALSE, warning=FALSE}
reasonstable<-fread("reasons.csv",sep=",")
reasonstable$Percentage <- (reasonstable$Respondents / 45)
reasonstable$Percentage <- percent(reasonstable$Percentage)


kable(reasonstable, caption = "Reasons for remote", col.names = c("Reason for remote team", "Respondents indicating this reason", "Percentage")) %>%
    row_spec(0, bold = T)
```

## Q: What seems to be more difficult when managing remote developers?

I standardized the free-form answers to this question using a simple coding system. The tabulated coded responses are below. Note that some respondents provided more than difficulty in managing remote developers.

```{r echo = FALSE, warning=FALSE}
difficultiestable<-fread("difficulty.csv",sep=",")
difficultiestable$Percentage <- (difficultiestable$Respondents / 45)
difficultiestable$Percentage <- percent(difficultiestable$Percentage)

kable(difficultiestable, caption = "Difficulties with remote", col.names = c("Primary difficulty", "Respondents indicating this difficulty", "Percentage")) %>%
    row_spec(0, bold = T)
```

# Insights

TODO

The median team size is 7. This gives the opportunity to see if there are "reasons why" and "main difficulty" differences between small and large teams.

Another slice: look at reasons why and see if that has any relationship to team size or problems.

# Method

In May 2019, I surveyed a group of Tech Leads and VP's of Engineering from Computer Software, Internet, and Information Technology and Services companies. This group was a convenience opt-in sample recruited using LinkedIn. I screened respondents by asking "In the last year, have you managed a dev team that is overseas OR a remote/distributed team that has just 2 or 3 hours/day of time zone overlap with you?"

- 56 people from LinkedIn completed the survey. 45 (80%) moved beyond the screener question.

# Open data

You can inspect and use for your own purposes the anonymized survey responses generated by this research: <TODO>

# Discussion

I welcome discussion with you about the conclusions you might draw from this data, especially if they differ from mine. 

# Brought To You By Philip Morgan Consulting

I'm Philip Morgan, and I'm focused on helping dev shops cultivate, commercialize, and monetize expertise. 

There's an underlying tension inherent in this expertise project: what dev shops need to do to cultivate exceptionally valuable expertise is often in direct conflict with what they need to do to serve clients -- and generate revenue -- today. My work helps resolve this tension. 

Learn more: 
<https://philipmorganconsulting.com/research/>

Template stuff:

`r tufte::margin_note('MARGINNOTE WITHOUT FOOTNOTE NUMBER')`
